Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2024-10-28 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
updated sweat test data for Stacey

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
library(forestmodel)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

>>>>

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

### 2024-10-28 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
updated sweat test data for Stacey

  PatientID, Diagnosis, Sex, Genotypes1, Genotypes2,
  Age.at.Test.in.Years, Method, Site, Na, K, Cl, Na_Left, Na_Right, K_Left,

@@

# Get each person's first test
chco <- chco %>%

@@

njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes.xlsx")

@@

@@

    `Patient ID`, Diagnosis, Sex, `Genotype 1`, `Genotype 2`,
    `Age at test (in years)`, `Test Method`, `Site of test`, `Sweat sodium`,

@@

@@

    subject, Diagnosis, sex, Genotypes1, Genotypes2, age, Method, site,

@@

# Limit to complete data

@@

    PatientID, Diagnosis, Sex, Age.at.Test.in.Years, Na:Cl,

@@

@@

write.csv(df, "./Data_Cleaned/analysis_dataset.csv", row.names = F, na = "")

@@

```{r}
#| message: false
#| warning: false
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(x = Na, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()

df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(
      x = Na, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
      group = Diagnosis
    )
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(
      x = Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
      group = Diagnosis
    )
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(
      x = Na / Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
      group = Diagnosis
    )
  ) +
  geom_point() +
  theme()
```

```{r}
df %>%
  filter(Diagnosis == "CF") %>%
  mutate(
    `Age group` = cut(Age.at.Test.in.Years, c(-Inf, 18, Inf), right = F),
    `Cl < 60` = Cl < 60
  ) %>%
  select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
  tbl_summary(by = `Age group`, missing_text = "Missing")
write.csv(df, file = "~/sweat_analysis_data.csv", na = "", row.names = F)
```

Add CRMS change in sweat and change in ratio

Add CF carrier genotypes to this report

Plot just the CF genotype function
  - Are the chloride values matching up with the CFTR function as we'd expect?

>>>>

  PatientID, Site, Diagnosis, Sex, Genotypes1, Genotypes2,
  Age.at.Test.in.Years, Method, Na, K, Cl, Na_Left, Na_Right, K_Left,

@@

@@

njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")

@@

  mutate(Site = "NJH") %>%

@@

    `Patient ID`, `Site of test`, Diagnosis, Sex, `Genotype 1`, `Genotype 2`,
    `Age at test (in years)`, `Test Method`, `Sweat sodium`,

@@

# Less detailed diagnoses
diagnoses$Diagnosis <- factor(diagnoses$Diagnosis,
  levels = c(
    "Abnormal sweat choride test without diagnosis of cystic fibrosis",
    "Abnormal sweat test",
    "Alpha-1-antitrypsin deficiency",
    "Alpha-1-antitrypsin deficiency carrier",
    "Alpha-1-antitrypsin deficiency, Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "CF", "CF carrier, Alpha-1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency carrier",
    "CF, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency carrier",
    "Cystic fibrosis screening",
    "Cystic fibrosis transmembrane conductance regulator (CFTR)-related disorder",
    "Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "Primary ciliary dyskinesia",
    "Primary Ciliary dyskinesia, Cystic Fibrosis carrier"
  ),
  labels = c(
    "Non-CF", "Non-CF", "Non-CF", "CF Carrier", "Non-CF", "Cystic Fibrosis",
    "CF Carrier", "Cystic Fibrosis", "Cystic Fibrosis", "Cystic Fibrosis",
    "CF Carrier", "CF Carrier", "CF Carrier", "Non-CF", "CFTR-RD or CRMS",
    "Non-CF", "Non-CF", "CF Carrier"

@@

    subject, site, Diagnosis, sex, Genotypes1, Genotypes2, age, Method,

@@

# Get rid of unnecessary columns

@@

    PatientID, Site, Diagnosis, Sex, Age.at.Test.in.Years, Na:Cl,

@@

# For each ID, get the row with the most data (or first test)
df$Missing <- rowSums(is.na(df))
  slice_min(Missing) %>%
  slice_min(row_number()) %>%
  filter(!is.na(Diagnosis)) %>%
  ungroup()

@@

write.csv(df, "./Data_Cleaned/final_analysis_dataset.csv", row.names = F, na = "")

@@

```{r}
#| message: false
#| warning: false
df %>%
  ggplot(
    aes(x = Na, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  ggplot(
    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  ggplot(
    aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
```

```{r}
df %>%
  filter(Diagnosis == "CF") %>%
  mutate(
    `Age group` = cut(Age.at.Test.in.Years, c(-Inf, 18, Inf), right = F),
    `Cl < 60` = Cl < 60
  ) %>%
  select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
  tbl_summary(by = `Age group`, missing_text = "Missing") %>%
  add_p()
```
